### 2018-06-06 Writing/response_to_reviewers.tex

\subsection{Major suggestions}

\begin{enumerate}

\item{\textcolor{blue}{I have some concerns about the approach, however. I was somewhat confused by the CD metric. When quantifying n, how is it decided how many taxa have "comparable" characters?
If you simulated the character for all taxa, do they not all have this character (i.e. n always equals the dataset size)?
But more importantly, will two characters not be more similar if they share the same number of character states, independent of if they are telling you the same evolutionary history?
It seems to me that whether or not two characters share a state space is not the relevant axis along which to measure correlation.}}

%TG: Specify the Gower distance (comparable)
%TG: yes, n is fixed here (this is just for the CD definition)
%TG: no, characters with missing data should be more dissimilar: since the similarity is measured in terms of shared splits, a missing data can always involve a new split (if the ? is a new state)
%TG: and not sure about the directionality thing.

%TG: TODO: clarify the CD metric definition

\item{\textcolor{blue}{I also found the explanations confusing for what proportion of a matrix was repeat sampled. For example, in a matrix that you call maximised, n characters are replaced. But how many is that really? What is the range? I had a hard time putting a context to this and deciding how well I feel these simulations approximate reality.  Without knowing this, I was having difficulty understanding why the statement "when character correlation was high (low character differences), the topology was always the furthest away from the``normal'' topology." 
should be true - does this simply mean that the same signal is being repeated over and over again, but that signal is low information relative to the topology? To my mind, highly correlated datasets would yield congruent trees, but that these trees may be discordant with the true tree, if the resampled characters were not representative of the true signal, or highly concordant with the true tree if they were. Perhaps this not being observed is a result of not performing comparisons to the true tree?}}

%TG: TODO: add the range of replaced character n and m
%TG: TODO: add more explanantion on the replacements in the methods and the discussion (the reviewer's thoughts)

@@

Following these two points, the reviewer counter example $CD_{(x,y)} = 0$ means that 1) $x'$ and $y'$ are identical and the characters $x$ and $y$ imply the same number of splits in a tree (here no splits since only one taxon is used - however, this would hold for any $n$ number of taxa).
We have clarified these two points in the manuscript and in the mathematical demonstration.

%TG: TODO: add the stuff to the manuscript

\item{\textcolor{blue}{While some details of the simulation study are included, some important ones are missing. In particular, in the design of the study, how did you avoid or mitigate errors introduced by the programs simulating data? All of the programs use random-number generators and have their own biases. How have you accounted for this in the study of design to conclude that your results are due to signal, and not bias from the programs used?}}

%TG: This is a tricky one. Simulations where generated using a random seed but each step of the simulation cascades down from it. In this context it is hard to control for how the random data was generated (seeds where used at the start of the matrix simulations and the start of the tree inferences to mitigate the effect).
%TG: In the end this is always a problem with computers and I'm not sure how to solve them unless using true random generators (overkill?).

\item{\textcolor{blue}{How did you ``randomly'' modify your matrices? Did you sample under some distribution? If so, which one?}}

%TG: TODO: add to the methods the sampling distribution (runif(0,1)).

@@

\item{\textcolor{blue}{It is not obvious why generated character sequences are representative of biological sequences or sample the space overall. Were the parameters chosen for the HKY model to fit a particular evolutionary scenario? Are they meant to capture all of the space?}}

%TG: again, not sure what the reviewer means, the HKY picks the HKY scenario and the Mk the GTR one.
%TG: and no, probably they don't capture all the space, morphological characters can be in any scenario (see Wright et al) but most of them seem to be GTR friendly.

@@

\item{\textcolor{blue}{Similarly, why did you choose 0.85 for the Mk model? Does it represent some important set of character sequences? How does it represent the space in general?}}

%TG: the 0.85 is the proportion of binary characters. This was extracted from litterature.

\item{\textcolor{blue}{There are statements that seem contradictory, like (line 335): ``Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability > 0.95 and different if they had an overlap probability > 0.05.''}}

We fixed this type. We considered them different if they had an overlap probability $<$ 0.05.

@@

\item{\textcolor{blue}{It is not clear how the conclusions were reached. For example, line 381-384: ``Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (Table 5 - median NTSRF of 0:828 and 0:679 respectively) but not in terms of individual taxon placements (Table 5 - median NTSTr of 0:738 and 0:601 respectively).'' Why is the 0:828 - 0:679 = 0:149 significant for this test and 0:738 - 0:601 = 0:137 is not?}}

The significances was referring to the results in Table 5 (Wilcox test p value of 0 and 0.08 and Bhattacharrya Coefficient of 0.891 and 0.984 respectively) and the median values were extracted from the supplementary material 3 Table 4.
We now detailed the values and where they are present for each of the tests to avoid confusion.

@@

\item{\textcolor{blue}{31:487 – and the use of models that utilise tree structure}}

We fixed this typo.

\item{\textcolor{blue}{32:492 – worth $->$ worst}}

We removed this sentence (see Reviewer 1 specific suggestion \ref{dishonest}).

### 2018-06-06 Writing/draft_highlighted.tex
Specified results locations

Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one.
The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see online material \url{https://rawgit.com/TGuillerme/CharactersCorrelation/master/Analysis/02-EffectCorrelationFullResults.html}). 
Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g. when multiple trees are equal to the ``normal'' tree).
Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146,soto2016trace}.
Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $>0.05$.
Comparisons falling between these range can not be designated as strictly similar/different but can still be ranked (e.g. for three distributions A, B, C, if $BC_{(A,B)} = 0.15$ and $BC_{(A,C)} = 0.65$, we cannot consider either distribution significantly different or similar but $B$ still has a lower probability of being similar to $A$ than $C$).

@@

The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303 - see supplementary material 3 for the full summary statistics).
There is a significant difference between all scenarios (maximising, minimising and randomising) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient of 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (Bhattacharrya coefficient of 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$ - Table \ref{Tab_pooledscenarios_test})

@@

\subsubsection{Number of characters}
This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively - see supplementary materials 3) with a significant difference only between 100 and 1000 and 350 and 1000 characters (Table \ref{Tab_pooledscharacters_test}).
The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3) with a significant difference between 100 and 350 or 1000 characters (Table \ref{Tab_pooledscharacters_test}).
However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to \hl{very} similar distributions.

@@

Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ and $NTS_{Tr}$ but no differences between 75 and 150 taxa (medians for 25, 75 and 150 taxa equals 0.802, 0.76, 0.763 $NTS_{RF}$ and 0.758, 0.588 and 0.615 $NTS_{Tr}$ respectively - Table \ref{Tab_pooledstaxa_test} and see supplementary materials 3).
Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).

@@

Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.828 and 0.679 respectively) but not in terms of individual taxon placements (Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.738 and 0.601 respectively).

>>>>

Because both $NTS_{RF}$ and $NTS_{Tr}$ metrics are bounded at one.
The residuals of any model based on the $NTS$ scores were not normal thus preventing the use of parametric tests for comparisons (see online material \url{https://rawgit.com/TGuillerme/CharactersCorrelation/master/Analysis/02-EffectCorrelationFullResults.html}). 
Similarly, a non-parametric Wilcoxon rank test \citep{hollander2013nonparametric} would be biased in its p-value calculation due to the presence of equal values in the $NTS$ distributions (e.g. when multiple trees are equal to the ``normal'' tree).
Therefore, we used a combination of the Wilcoxon rank test with a Bonferonni-Holm corrections \citep[to ensure our significant results were robust to Type I error rate inflation;][]{holm1979simple} and a simple non-parametric metric for measuring the probability of overlap between two distributions, the Bhattacharyya Coefficient \citep[$BC$;][]{Bhattacharyya,Guillerme2016146,soto2016trace}.
Thus, additionally to the Wilcoxon test results, we considered distribution to be significantly similar if they had an overlap probability $>0.95$ and different if they had an overlap probability $<0.05$.
Comparisons falling between these range can not be designated as strictly similar/different but can still be ranked (e.g. for three distributions A, B, C, if $BC_{(A,B)} = 0.15$ and $BC_{(A,C)} = 0.65$, we cannot consider either distribution significantly different or similar but $B$ still has a lower probability of being similar to $A$ than $C$).

@@

The overall amount of character difference in a matrix has an effect of the ability to recover the correct topology when maximising character difference leading to the smallest loss in phylogenetic information (median $NTS_{RF}$ = 0.956 and median $NTS_{Tr}$ = 0.839\hl{; see supplementary materials 3, Table 1}) followed by simply randomising the characters (median $NTS_{RF}$ = 0.762 and median $NTS_{Tr}$ = 0.628\hl{; see supplementary materials 3, Table 1}) and minimising the character difference (median $NTS_{RF}$ = 0.605 and median $NTS_{Tr}$ = 0.303\hl{; see supplementary materials 3, Table 1}).
There is a significant difference between all scenarios (maximising, minimising and randomising; \hl{\textit{p} = 0; Table} \ref{Tab_pooledscenarios_test}) with the highest probability of overlap being between maximising and randomising the character difference (Bhattacharrya Coefficient (BC) = 0.873 for the $NTS_{RF}$ and 0.908 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test}) and the lowest probability between maximising and minimising the character difference (BC = 0.573 for the $NTS_{RF}$ and 0.614 for the $NTS_{Tr}$; Table \ref{Tab_pooledscenarios_test})

@@

\subsubsection{Number of characters}
This effect of the character difference is not dependent on the number of characters when looking at clade conservation (i.e. $NTS_{RF}$).
The median $NTS_{RF}$ was similar for 100, 350 and 1000 characters (0.730, 0.745, 0.767 respectively; see supplementary materials 3 \hl{Table 2}) with a significant difference only between 100 and 1000 and 350 and 1000 characters (\hl{\textit{p} = 0.001 and \textit{p} = 0.032 respectively}; Table \ref{Tab_pooledscharacters_test}).
The number of characters affects the character difference more in terms of taxon placement for a low number of characters (median $NTS_{Tr}$ for 100, 350 and 1000 characters equals 0.544, 0.693, 0.799 respectively - see supplementary materials 3 \hl{Table 2}) with a significant difference between 100 and 350 or 1000 characters (\hl{\textit{p} = 0, \textit{p} = 0 and \textit{p} = 0.014 respectively}; Table \ref{Tab_pooledscharacters_test}).
However, these differences have to be contrasted by a very high probability of overlap between each number of characters and metrics (Bhattacharrya Coefficient always $>0.95$) suggesting that the significant effects of the number of characters still leads to \hl{very} similar distributions.

@@

Similar to the effect of number of characters on character difference, the number of taxa seems to have only a marginal effect.
A low number of taxa (25) resulted in significant differences with both 75 or 150 taxa in both $NTS_{RF}$ and $NTS_{Tr}$ but no differences between 75 and 150 taxa (\hl{\textit{p} = 0.012, \textit{p} = 0.004, \textit{p} = 0 and \textit{p} = 0 respectively}; Table \ref{Tab_pooledstaxa_test} and medians for 25, 75 and 150 taxa equals 0.802, 0.76, 0.763 $NTS_{RF}$ and 0.758, 0.588 and 0.615 $NTS_{Tr}$ respectively; see supplementary materials 3 \hl{Table 3}).
Again, however, the significant differences have to be contrasted with still high probabilities of overlaps for each $NTS_{RF}$ and $NTS_{Tr}$ distributions for every number of taxa (Table \ref{Tab_pooledstaxa_test}).

@@

Regarding the inference method, there is a significant difference in clade conservation between Bayesian and maximum parsimony (\hl{\textit{p} = 0} Table \ref{Tab_pooledsmethods_test} - median $NTS_{RF}$ of 0.828 and 0.679 respectively; \hl{see supplementary materials 3, Table 4}) but not in terms of individual taxon placements (\hl{\textit{p} = 0.08}; Table \ref{Tab_pooledsmethods_test} - median $NTS_{Tr}$ of 0.738 and 0.601 respectively; \hl{see supplementary materials 3, Table 4}).
